Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes

### 2020-05-14 Ben_Manuscripts/hdphmm/Draft.tex
New supporting figures and mods to draft

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{amsmath} % or simply amstext
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\newcommand{\angstrom}{\textup{\AA}}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}
\usepackage{xr}
%\DeclareUnicodeCharacter{03C9}{$\omega$}
\externaldocument[S-]{Supporting_Information}

@@

  To cluster on $T_{ii}$, we first cast it in terms of the expected dwell times. The expected
  value of the number of sequential self-transitions is simply $\frac{1}{1 - T_{ii}}$. This 
  relationship correctly implies that dwell times approach infinity as $T_{ii}$ approaches
  1. Therefore we found the most success with agglomerative clustering by linearizing this
  relationship and clustering on $-\log(1 - T_{ii})$.
  % (see Section~\ref{S-section:selfT_cluster} of the Supporting Information for a graphical illustration).

@@

  \textit{Generating Stochastic IHMM Trajectory Realizations}: For each independent 
  MD trajectory, we generated stochastic trajectory realizations using the finalized 
  parameters of the associated IHMM. We drew state sequences with transition probabilities
  given by $T$. While in a given state, we simulated motion according to the VAR(1) 
  parameterization of that state. After each state transition, we set the unconditional
  mean of the new state to the particle's position immediately before the state 
  transition occurred.

  \textit{Calculating MSD}: We quantified solute motion using the time averaged mean squared displacement (MSD)
  of the solute center of mass trajectories.~\cite{meroz_toolbox_2015} The time-averaged MSD measures all observed
  displacements over time lag $\tau$:
  \begin{equation}
  	\overline{z^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (z(t + \tau) - z(t))^2 dt
  \label{eqn:tamsd}
  \end{equation}
  where T is the length of the trajectory. We reported $1 \sigma$ confidence intervals 
  based on the results of 200 bootstrap trials. For each trail, we calculated the mean of 
  $n$ random trajectories chosen with replacement from the pool of trajectories, where 
  $n$ is the number of trajectories.

  The MSD reported for each solute measured by our MD 
  simulations is a distribution created from the 24 independent MSD curves. We predicted 
  MSDs and their uncertainties with our stochastic models by bootstrapping the 
  distribution of stochastically generated MSDs. For each bootstrap trial, we generated
  24 independent trajectories using the 24 sets of IHMM parameters. We report the mean
  and standard error of the bootstrapped MSD distributions.

@@

  We determined the number of hydrogen bonds between each solute and the membrane
  as a function of time. Based on the geometric criteria of Luzar and Chandler, we 
  define a hydrogen bond to exist if the distance between donor, D, and acceptor, 
  A, atoms is less than 3.5\AA~and the angle formed by $D-H \cdots A$ is less than 
  $30\degree$.~\cite{luzar_effect_1996}

  We estimated the lifetime of hydrogen bonds by recording the length of 
  sequential frames where solutes remained hydrogen bonded. We still counted sequences
  where hydrogen bonds were broken for a single frame before reforming. Consistent
  with our previous work, we reported the 95th percentile of hydrogen bond lifetimes
  since their distribution is not Gaussian and to emphasize longer trapping periods.~\cite{coscia_chemically_2019}

  We also measured the degree of association between solutes and sodium ions on a
  frame-by-frame basis. We define a sodium ion to be associated with an atom if they
  are within 2.5 \AA~of each other, as determined in our previous work.~\cite{coscia_chemically_2019}

  %We developed a way to 
  We measure the local density of the membrane at arbitrary
  points in the unit cell as follows. We histogrammed the positions in three dimensions
  Since our system is in a monoclinic unit cell, we periodically replicated
  the system in the $\pm x$, $y$ and $z$ directions and then chose the bounds on the
  histogram in order to create a rectangular box encompassing the unit cell with
  a 1 nm buffer between the histogram and unit cell boundaries.~\cite{van_der_walt_numpy_2011} We then used a regular
  grid interpolator in order to allow interpolation at arbitrary points within the grid.~\cite{virtanen_scipy_2020}

@@

  As shown in Figure~\ref{fig:qualitative_unclustered}, realizations of our model 
  can produce trajectories that show qualitatively similar hopping and trapping 
  behavior to the MD trajectories. It is important that our model reproduces this 
  behavior so that we can have confidence in its usefulness for quantitative 
  predictions.

  % BJC4: This figure needs to be redone with new trajectory generation procedure
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{qualitative_unclustered_MET2.pdf}
  \caption{We can generate trajectories which bear qualitative resemblance to
  the MD trajectories. Solute trajectories generated by the IHMM (blue) show
  the same hopping and trapping behavior exhibited by MD (black).
  }\label{fig:qualitative_unclustered}
  \end{figure}

  \textit{Deviations of the IHMM from Molecular Motion}: From a quantitative
  standpoint, using the IHMM as a predictive tool has two 
  shortcomings. First, in many cases, the MSD predicted by realizations of the 
  IHMM under-predicts that of MD (see Figure~\ref{fig:unclustered_msds}).
  Second, the shapes of the MSD curves at short time lags are very different.

@@

  The IHMM under-predicts the MD MSD due to its tendency to favor fewer states. 
  Disjoint segments with similar VAR parameters, but different lengths of entrapment
  can be parameterized as the same state. In Figure~\ref{fig:dwell_distributions}, 
  we compare the overall dwell time distributions generated directly from MD simulations to that 
  generated from realizations of the IHMM. We do this by measuring the time spent in 
  each independent same-state segment. Some of the probability density of short 
  trapping times in the MD simulations gets shifted to medium length trapping times
  in realizations of the IHMM. In Figure~\ref{fig:dwell_distributions_modT} 
  and~\ref{fig:msd_modT}, we show that by modifying the transition matrix to 
  encourage transitions towards states with shorter dwell times, we can more closely
  reproduce the dwell time distribution and MSD of MD. This is because most solute
  motion is a consequence of transitions between trapped states.

  % BJC4: A lower priority supporting information section I've had in mind is one
  % with tips for fitting and validating the IHMM on individual trajectories.  
%  In many cases, especially evident where MSD predictions over-estimate MD, there
%  is some error in the state sequence segmentation. In Section~\ref{S-section:fitting_tips}
%  of the Supporting Information, we give some tips based on our experience of how to 
%  initialize IHMM fits in order to reduce error.

  \begin{figure}
  \centering
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{dwell_distributions.pdf}
  \caption{}\label{fig:dwell_distributions}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{dwell_distributions_modT.pdf}
  \caption{}\label{fig:dwell_distributions_modT}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{unclustered_msd_MET_modT.pdf}
  \caption{}\label{fig:msd_modT}
  \end{subfigure}
  \caption{(a) The density of short dwell times compiled from realizations of
  the IHMM fit to methanol trajectories is below that measured directly from 
  the MD trajectories. If we modify the transition matrix in order to favor 
  transitions towards states with short dwell times, we can more closely match
  the (b) dwell time distributions and (c) MSDs.}\label{fig:slow_T}
  \end{figure}

  At extremely short time lags, the MSD is generally over-predicted by the IHMM
  (see Figure~\ref{fig:unclustered_msd_shortlag}) because the VAR(1) model assumes
  multivariate Gaussian noise while the MD data suggests the noise is better 
  modeled by more general L\'evy stable noise. The first time lag of the MSD is
  equivalent to the variance of all observed fluctuations (first-order difference). 
  The IHMM simulation appears to over-estimate the variance because the actual 
  distribution of solute fluctuations has heavy tails with a smaller variance. (Figure~\ref{fig:emission_widths}).
  The total distribution of hops is expected to be heavy tailed because it is a
  combination of hop distributions from many states with different variances. More
  interestingly, in Figure ~\ref{fig:state_emission_widths}, we show that the 
  emission distribution of a single state picked out by the IHMM is fit better by
  a more general L\'evy distribution while the emissions from IHMM realizations are
  Gaussian, as expected. An improvement to the IHMM that may allow dynamics more 
  faithful to MD would use an autoregressive model with L\'evy stable noise. However,
  this would be less computationally efficient due to lack of conjugate priors.

  \begin{figure}
  \centering
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{unclustered_msd_MET_shortlag.pdf}
  \caption{}\label{fig:unclustered_msd_shortlag}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{emission_widths.pdf}
  \caption{}\label{fig:emission_widths}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{state_emission_widths.pdf}
  \caption{}\label{fig:state_emission_widths}
  \end{subfigure}
  \caption{(a) At short time lags, the MD MSD (black) and the IHMM MSD (orange)
  differ significantly in their shape and curvature. (b) While the 
  fluctuations in $z$ of both distributions are heavy-tailed, fitting well to zero-centered
  symmetric L\'evy stable distributions with stability parameters, $\alpha$, the IHMM has
  a wider variance. (c) The distribution of fluctuations in a chosen state frequently
  visited by methanol is heavy tailed, but the IHMM is constrained to produce fluctuations
  from a multivariate normal distribution ($\alpha$=2 in each dimension).
  }\label{fig:short_timelags}
  \end{figure}

  Also apparent from Figure~\ref{fig:unclustered_msd_shortlag} is that the 
  curvature of the MD MSD is far more gradual than that generated by the IHMM.
  This is because the VAR(1) model is only correlated to its previous fluctuation
  while the MD simulations suggest that correlation persists for at least tens
  of nanoseconds.
    % BJC2: could probably use a correlation function to suggest the AR order and stick the plot in SI
  One may be able to reproduce the curvature at short time lags with higher order
  VAR models, but it becomes a much higher dimensional parameterization that would
  require significantly more data in order to produce stable results. Since we 
  are interested in reproducing longer time scale behavior, it is not an issue of
  great concern.

  % BJC2: I'm not sure if the width of the hop length distribution has a significant 
  % effect on average MSD, since it's symmetric. Something to test next week.

@@

  % BJC2: I think it makes sense to talk about selectivity here, but I'm not sure 
  % the best way to go about it.

  Keeping in mind the the shortcomings of the IHMM, we can now propagate realizations onto long 
  time scales in order to predict solute flux and thus selectivity. 
  \begin{itemize}
    \item In our previous work, we showed that in systems without convective flux, selectivity is
    simply the ratio of solute fluxes.
%    \item Although the magnitude of our MSD predictions are typically below those of
%    MD, we may still obtain reasonable estimates of selectivity.
  \end{itemize}

  We can use the ratio of solute fluxes predicted by realizations of our models in order to 
  estimate selectivity.
  \begin{itemize}
    \item Exact methodology under development
    \item Will probably just take advantage of long timescale MSD linearity in order to get
    flux, rather than mean first passage time approach.
    \item Need to figure out how to deal with differences in MSD under-estimates
  \end{itemize}

@@

  \subsubsection*{State Clustering}

  In order to reduce the state space to a more manageable size, and to allow
  different solute trajectories to share similar behavior, we clustered the
  states identified by the IHMM for each solute as described in 
  Section~\ref{method:clustering}. In Figure~\ref{fig:mechanism_map}, we show
  the results of clustering on the same trajectory from 
  Figure~\ref{fig:rz_unclustered}. We chose to use X clusters based on the
  criteria described in Section~\ref{method:clustering}. Therefore, we reduced 
  the total state space from 300 to X. By working to understand the parameters
  of each cluster, we can gain an understanding of associated solute transport
  mechanisms responsible for those parameters.

  \subsubsection*{Relating Clusters to Mechanisms}

  The parameters of each cluster can be related to dominant solute-membrane interactions.
  \begin{itemize}
    \item Based on our previous work, we know that hydrogen bonding, sodium ion 
    association and radial location within the membrane pores will have a significant
    influence on transport properties. 
    \item Even without our previous work, it would still be easy to speculate that these 
    interactions might occur and could therefore be used to help understand the 
    states identified by the IHMM.
    \item We have also incorporated membrane density as a physical metric because we 
    know the membrane has an inhomogeneous structure.
  \end{itemize}

  \textit{Learning from individual trajectories}: In Figure~\ref{fig:mechanism_map}, 
  we demonstrate the wide range of behavior exhibited by a representative 2D methanol
  trajectory.
%BJC2: reducing duplication
%  The coordinates
%  are colored according to the cluster with which methanol's dynamics are most consistent,
%  which implies that this particular methanol trajectory exhibits eight distinct dynamical 
%  behaviors (in order of their first appearance, they are colored dark yellow, red, light
%  blue, light green, orange, blue, teal and lime green).
  The solute time series are coupled with color coded bars which describe the solute's
  local environment and physical interactions that it undergoes at each trajectory frame. 
%  From the top, they describe 
%  whether the solute is associated with sodium, whether the solute is hydrogen bonding
%  with the monomer as well as to what part of the monomer it is hydrogen bonding, and 
%  finally, the local number density surrounding the solute.

  %BJC2: the exact figure will change, but most of what I say shouldn't
  \begin{figure}
  \centering
  \begin{subfigure}{0.75\textwidth}
  \includegraphics[width=\textwidth]{mechanism_map.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}{0.24\textwidth}
  \includegraphics[width=\textwidth]{monomer_oxygens.png}
  \caption{}
  \end{subfigure}
  %MRS: a lot of this caption test is duplicated from the main text. I would suggest reducing the duplication.
  %BJC: you think better to have more detail in text or in caption?
  %MRS2: at a certain point, it should be in the text.  In the caption, I would just state what the point of the plot is, 
  % and define the colors/data/etc.  But full explanation would be better in the text. Captions should be scannable quickly,
  % based on how people process papers, and this should be significantly shorter for readability. 
 \caption{(a) We can learn a significant amount of detail about solute motion by viewing 
  solute trajectories color-coded according to their clustered dynamical behavior alongside
  plots of the dominant molecular interactions and descriptions of the membrane's local 
  environment. In the plot above, we show the radial and axial coordinates of a single
  methanol molecule over the course of the equilibrated portion of its MD simulation.
  Disjoint segments of the same color imply the solute behaves similarly at different 
  points of the trajectory because their VAR parameters place them in the same cluster. 
  There are eight distinct dynamical modes exhibited by this trajectory. In order
  of their first appearance, they are colored dark yellow, red, light blue, light green,
  orange, blue, teal and lime green. On the top of the plot are bars which represent different
  physical interactions and the membrane's local environment as a function of time. The top 
  bar is colored green at time points where solutes are associated with sodium. Methanol 
  associates with sodium ions relatively infrequently. The next bar down is colored when hydrogen
  bonds exist. The colors match the enlarged oxygen atoms highlighted in (b). Blue slivers
  correspond to hydrogen bonds donated to the monomer's carboxylate head group. Green slivers
  correspond to  hydrogen bonds donated to the ether linkages between the head groups
  and tails. Red slivers correspond to hydrogen bonds donated to the oxygens on the ends
  of the monomer tails. Methanol exhibits significant hydrogen bonding with all parts of
  the monomer. Which part of the monomer methanol hydrogen bonds to is heavily correlated
  to the solute's $r$ coordinate. Finally, the bar with the blue gradient represents the local 
  membrane density based on the solute's ($r$, $z$) coordinate. Darker shades of blue
  correspond to higher membrane densities. The density appears to be higher when methanol
  motion is restricted (e.g. from 75 to 550 ns). It is relatively low in areas where methanol's
  position fluctuates significantly (e.g. from 550 to 700 ns). 
  }\label{fig:mechanism_map}
  \end{figure}

  %MRS2: suggest subsubcaptions/italics at the start of parentheses to make it easier for a reader to see the structure.
  %MRS2: like \textit{hydrogen bonding patterns}: etc.

  One can gain useful chemical intuition just by studying plots like 
  Figure~\ref{fig:mechanism_map}. It is clear that, for methanol, association 
  with sodium is a much rarer interaction than hydrogen bonding with the 
  LLC monomers. Methanol hydrogen bonds to all regions of the monomers dependent
  on its radial position. Its fluctuations tend to be smallest when hydrogen bonded
  or in areas of high local number density. For example, from approximately 75 to 
  525 ns, methanol appears to be trapped in a high density region of the tails.
  In the time that follows, it enters a region of relatively low density where 
  fluctuations are quite large. During this time period, there is intermittent, 
  short-lived hydrogen bonding with the oxygen atoms at the tail ends, as suggested
  by the frequent state changes during that time period and red slivers in the bar
  above. Significant hops in the $z$ direction appear to be described by the dark blue
  state. After entering the dark blue state ca. 850 ns, methanol appears to jump
  significantly in the $z$ direction from a hydrogen bonded state with the ether 
  linkages to a hydrogen bonded state with a head group in a monomer below it.

  %MRS2: not entirely clear about the purpose of this paragraph? Seems to be just saying it's possible to learn more?
  %MRS2: should clarify.
  %BJC3: It was meant to be transitory but can probably cut out with the addition of the subsection title
%  It is clear that there is a wealth of information in these plots, but they don't give
%  the complete picture needed in order to formulate LLC membrane design principles. One
%  can analyze each segment sequentially and form a clear picture of this instance of 
%  methanol's behavior without the need to visualize the trajectory with software like VMD.~\cite{humphrey_vmd:_1996}
%  However, it is still possible to gain a more general understanding of methanol's average
%  behavior.

  % BJC2: the numbers and choices of states in this paragraph will change, but I think we
  % can still use this framework to study mechanisms.
  \textit{Learning from all trajectories}: We may learn the most by studying the
  dynamical modes common to the majority of 
  the solute trajectories. Using the IHMM, we identified 33 total dynamical modes 
  exhibited by methanol. Of those, only six states appear in a third (8 out of 24) 
  or more methanol trajectories, and three of them appear in over 50 \% of the 
  methanol trajectories. 12 of the detected states are shared by two or less 
  methanol trajectories. In Figure~\ref{fig:common_states_MET_lines}, we plot 
  representative dynamics of each of the six states which we determined to be prevalent.
  We also chose to study a seventh state, which only appears in three out of
  24 trajectories, because methanol experiences very long periods of entrapment
  while in this state. Long trapping events, even if they are relatively rare, will
  have a significant influence on solute flux. No trajectory contains all seven states.
  However, the trajectory in Figure~\ref{fig:mechanism_map} exhibits all modes except
  states 6 and 7. The states in Figure~\ref{fig:common_states_MET} are colored to match
  states that also appear in Figure~\ref{fig:mechanism_map}.

>>>>

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{amsmath} % or simply amstext
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\newcommand{\angstrom}{\textup{\AA}}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}
\usepackage{xr}
\usepackage{enumitem}
\newcommand{\nclusters}{10}
%\DeclareUnicodeCharacter{03C9}{$\omega$}
\externaldocument[S-]{Supporting_Information}

@@

  To cluster on $T_{ii}$, we first cast it in terms of the expected dwell times. The expected
  value of the number of sequential self-transitions is simply:
  % BJC4: I numbered this equation because I reference it later.
  \begin{equation}  
   E[dwell time] = \frac{1}{1 - T_{ii}}
   \label{eqn:dwell_times}
  \end{equation}  
  This relationship implies that dwell times approach infinity as $T_{ii}$ approaches
  1. Therefore we found the most success with agglomerative clustering by linearizing this
  relationship and clustering on $-\log(1 - T_{ii})$.

@@

  \textit{Generating Stochastic IHMM Trajectory Realizations}: We used two 
  different methods to stochastically generate trajectories:
  \begin{enumerate}[label={Method \theenumi :}, leftmargin=3.5\parindent]
    \item For each independent MD trajectory, we generated stochastic trajectory
    realizations using the finalized parameters of the associated IHMM. This 
    results in 24 separate ensembles of trajectories for each solute studied. 
    \item Using the clustered parameter sets, we generated a single ensemble
	of trajectories. Realizations are representative of the combined behavior
	of all solute trajectories.
  \end{enumerate}

  For both approaches, we drew state sequences with transition probabilities given 
  by $T$ of the IHMM. While in a given state, we simulated motion according to the 
  VAR(1) parameterization of that state. After each state transition, we set the 
  unconditional mean of the new state to the particle's position immediately before
  the state transition occurred.

  \textit{Calculating mean squared displacement}: We quantified solute motion from
  MD and from our stochastic trajectories using the time averaged mean squared 
  displacement (MSD) of the axial ($z$) component of the solute center of mass
  trajectories.~\cite{meroz_toolbox_2015} The time-averaged MSD measures all observed 
  displacements over time lag $\tau$:
  \begin{equation}
  	\overline{z^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (z(t + \tau) - z(t))^2 dt
  \label{eqn:tamsd}
  \end{equation}
  where T is the length of the trajectory.

  We reported the mean and $1 \sigma$ confidence intervals of the MSDs based on the 
  results of 200 bootstrap trials. Our bootstrapping procedure varied dependent on the
  method of trajectory generation:
  \begin{enumerate}[label={Method \theenumi :}, leftmargin=3.5\parindent]
  	\item For each trial, we generated 24 stochastic trajectories, one from each of
  	the 24 independent IHMM models. The result of the bootstrap trial is the mean
  	of the MSDs calculated from each trajectory. 
    \item We generated an ensemble of 1000 independent stochastic trajectories using
    the clustered parameter sets. For each trial, we calculated the MSD of 1000 
    trajectories sampled, with replacement, from the ensemble.
  \end{enumerate}

  For MD trajectories, we bootstrapped according to Method 2.

  \subsection{Predicting Selectivity}\label{method:selectivity}

  In our previous work, we showed that in systems without convective flux, such
  as our MD simulations, selectivity towards component $i$ over $j$ is simply:
  \begin{equation}
  S_{ij} = \frac{J_i / \Delta C_i}{J_j / \Delta C_j}
  \label{eqn:selectivity_flux}
  \end{equation}
  where $J$ is solute flux and $\Delta C$ is the trans-membrane concentration
  difference.~\cite{coscia_capturing_2020} The concentrations of solutes are the
  same across our simulations, therefore Equation~\ref{eqn:selectivity_flux} 
  reduces to the ratio of solute fluxes.

  Solute flux, as written, can be expressed in terms of the product of solute 
  diffusivity, $D_i$, within the membrane and a sorption coefficient, $K_i$, 
  that quantifies the ability of the solute to move between the bulk solution
  and and the membrane:
  \begin{equation}
    J_i = \frac{D_i K_i}{l}\Delta C
  \end{equation}
  where $l$ is the membrane thickness. For our purposes, $K_i$ and $l$ are 
  irrelevant since we do not study pore entry effects or flux across finite
  length pores. Therefore, in order to calculate selectivity, with respect
  to quantities we can study with our MD simulations, it is only necessary that
  we compare the ratio of diffusion constants:
  \begin{equation}
    S_{ij} = \frac{D_i}{D_j}
    \label{eqn:selectivity_diffusivity}
  \end{equation}
  It can be shown on time scales where motion is uncorrelated, that the diffusion
  constant is linearly proportional to the slope of the MSD curve:~\cite{einstein_investigations_1956}
  \begin{equation}
    \overline{z^2(\tau)} = 2 n D t
    \label{eqn:msd_D}
  \end{equation}
  where $n$ is the number of spatial dimensions (1 in our case). Since the MSD
  of our models enters a linear regime, we can estimate diffusion constants 
  based on the slope of the linear region of the predicted MSD curves.

@@

  \textit{Expected dwell times}:

  \textit{Hydrogen bond detection}: We determined the number of hydrogen bonds 
  between each solute and the membrane as a function of time. Based on the 
  geometric criteria of Luzar and Chandler, we define a hydrogen bond to exist 
  if the distance between donor, D, and acceptor, A, atoms is less than 
  3.5\AA~and the angle formed by $D-H \cdots A$ is less than $30\degree$.~\cite{luzar_effect_1996}

  \textit{Hydrogen bond lifetimes}: We estimated the lifetime of hydrogen bonds by 
  recording the length of sequential frames where solutes remained hydrogen bonded.
  We still counted sequences where hydrogen bonds were broken for a single frame 
  before reforming. Consistent with our previous work, we reported the 95th 
  percentile of hydrogen bond lifetimes since their distribution is not Gaussian
  and to emphasize longer trapping periods.~\cite{coscia_chemically_2019}

  \textit{Sodium ion association detection}: We also measured the degree of association
  between solutes and sodium ions on a frame-by-frame basis. We define a sodium ion to
  be associated with an atom if they are within 2.5 \AA~of each other, as determined 
  in our previous work.~\cite{coscia_chemically_2019}

  %We developed a way to 
  \textit{Local density measurements}: We measured the local density of the membrane at
  arbitrary points in the unit cell as follows. We histogrammed the positions in three 
  dimensions. Since our system is in a monoclinic unit cell, we periodically replicated
  the system in the $\pm x$, $y$ and $z$ directions and then chose the bounds on the
  histogram in order to create a rectangular box encompassing the unit cell with
  a 1 nm buffer between the histogram and unit cell boundaries.~\cite{van_der_walt_numpy_2011}
  We then used a regular grid interpolator in order to allow interpolation at arbitrary
  points within the grid.~\cite{virtanen_scipy_2020}

@@

  As shown in Figure~\ref{fig:qualitative_unclustered}, realizations of our model 
  can produce trajectories that show qualitatively similar hopping and trapping 
  behavior to the MD trajectories to which they were fit. It is important that our
  model reproduces this behavior so that we can have confidence in its usefulness
  for quantitative predictions.

  % BJC4: This figure needs to be redone with new trajectory generation procedure
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{qualitative_unclustered_MET2.pdf}
  \caption{We can use realizations of the IHMM to generate trajectories which 
  bear qualitative resemblance to the MD trajectories. We used the parameters 
  of the IHMM, whose state segments are visualized in 
  Figure~\ref{fig:rz_unclustered}, in order to construct an ensemble of 
  characteristic trajectories. Solute trajectories generated with
  the IHMM (blue) show the same hopping and trapping behavior exhibited by 
  MD (black).
  }\label{fig:qualitative_unclustered}
  \end{figure}

  In Figure~\ref{fig:unclustered_msds}, we show that we can reproduce the 
  MD MSDs on long timescales within uncertainty using realizations of the 
  IHMM.
  \begin{itemize}
    \item We generated the MSD curves using Method 1 described in 
    Section~\ref{method:realizations} of the methods.
    \item Our ability to reproduce these curves gives use confidence
    in their long term predictions.
  \end{itemize}

@@

@@

  Since we can generate trajectories that share features similar to MD simulations, 
  it is now reasonable to project solute behavior onto much longer time scales.
  \begin{itemize}
    \item The IHMM MSD predictions quickly enter a linear regime, therefore we can 
    project long timescale MSDs based on the predicted diffusion constant as described
    in Section~\ref{method:selectivity}.
  \end{itemize}

  %BJC4: figure here

  %BJC4: some kind of brief selectivity analysis here

@@

  \textit{State Clustering}: In order to reduce the state space to a more 
  manageable size, and to allow different solute trajectories to share 
  similar behavior, we clustered the states identified by the IHMM for each
  solute as described in Section~\ref{method:clustering}. In 
  Figure~\ref{fig:clustered_traj_MET}, we show the results of clustering on the 
  same methanol trajectory from Figure~\ref{fig:rz_unclustered}. Based on 
  the criteria described in Section~\ref{method:clustering},
  we chose to reducing the total state space for methanol from 287 to \nclusters.

  Our goal is to derive the simplest model which adequately distinguishes
  different types of solute motion. By working to understand the parameters of 
  each cluster, we can uncover the associated transport mechanisms responsible
  for those parameters.

  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{clustered_traj_MET_ward_10.pdf}
  \caption{We reduced the total state space from 287 to 10 total states 
  shared across all trajectories. For the trajectory in this figure, we reduced
  the total number of states from 24 (see Figure~\ref{fig:rz_unclustered}) to
  5.}\label{fig:clustered_traj_MET}
  \end{figure}

  \textit{Dynamics of clustered parameter sets}: The clustered parameter sets
  allow transitions between dynamics exhibited by the same solute from independent 
  trajectories. This allows us to create hybrid trajectories which draw on all
  solute behavior.

  Realizations based on the 10 clustered parameter sets do not show the same
  qualitative dynamics as the MD simulation trajectories. 
  \begin{itemize}  
    \item In Figure~\ref{fig:qualitative_clustered_MET}, we compare MD 
    trajectories to realizations of our model using the clustered parameter sets.
    \item They lack the diversity of behavior which we see in the MD simulation trajectories.
    \item They do not show the same range of trapping times and hop lengths.
    \item If we use a higher number of clusters, the qualitative match improves, but
    at the cost of a larger state space to interpret (see Figure~\ref{S-fig:qualitative_improvement}
    of the Supporting Information).
  \end{itemize}

  In addition to qualitative mismatches, realizations of the clustered IHMM 
  under-predict the unclustered MSD predictions and the MD MSD. 
  \begin{itemize}
    \item Much of a solute's MSD is likely a consequence of short-lived trapping states
    with high variances.
    \item When clustering, disjoint segments with similar enough parameters are
    re-parameterized together as part of the same state, lengthening the lifetimes 
    of very short-lived states and suppressing the fluctuations of high variance states.
    \item In Figure~\ref{fig:clustered_MSD}, it is evident that increasing the number
    of clusters increases the predicted MSD.
    \item Matching the unclustered MSDs would likely require many more clusters, at 
    which point the benefits of clustering are lost. 
    %BJC: I don't think the following discussion is necessary anymore.
%    \item In Figure~\ref{fig:dwell_distributions}, we compare the overall dwell time
%    distributions generated directly from MD simulations to that generated from 
%    realizations of the IHMM. 
%    \item We do this by measuring the time spent in each independent same-state segment.
%    \item Some of the probability density of short trapping times in the MD simulations 
%    gets shifted to medium length trapping times in realizations of the IHMM. 
%    \item In Figure~\ref{fig:dwell_distributions_modT} and~\ref{fig:msd_modT}, we show 
%    that by modifying the transition matrix to encourage transitions towards states with
%    shorter dwell times, we can more closely reproduce the dwell time distribution and 
%    MSD of MD. This is because most solute motion is a consequence of transitions between
%    trapped states.
  \end{itemize}

  \begin{figure}
  \centering
  \begin{subfigure}{0.63\textwidth}
  \includegraphics[width=\textwidth]{qualitative_clustered_MET_10.pdf}
  \caption{}\label{fig:qualitative_clustered_MET}
  \end{subfigure}
  \begin{subfigure}{0.35\textwidth}
  \vspace{2.25em}
  \includegraphics[width=\textwidth]{msd_nclusters.pdf}
  \caption{}\label{fig:clustered_MSD}
  \end{subfigure}
  \caption{(a) Stochastic trajectory realizations generated from our model (blue) based on
  the clustered methanol parameters do not qualit}\label{fig:clustered_dynamics}
  \end{figure}

%  \begin{figure}
%  \centering
%  % BJC4: combine (a) and (b). Add figure with MSD prediction
%  \begin{subfigure}{0.32\textwidth}
%  \includegraphics[width=\textwidth]{dwell_distributions.pdf}
%  \caption{}\label{fig:dwell_distributions}
%  \end{subfigure}
%  \begin{subfigure}{0.32\textwidth}
%  \includegraphics[width=\textwidth]{dwell_distributions_modT.pdf}
%  \caption{}\label{fig:dwell_distributions_modT}
%  \end{subfigure}
%  \begin{subfigure}{0.32\textwidth}
%  \includegraphics[width=\textwidth]{unclustered_msd_MET_modT.pdf}
%  \caption{}\label{fig:msd_modT}
%  \end{subfigure}
%  \caption{(a) The density of short dwell times compiled from realizations of
%  the IHMM fit to methanol trajectories is below that measured directly from 
%  the MD trajectories. If we modify the transition matrix in order to favor 
%  transitions towards states with short dwell times, we can more closely match
%  the (b) dwell time distributions and (c) MSDs.}\label{fig:slow_T}
%  \end{figure}

% BJC: probably a bad idea
%  It may still be possible to use these hybrid trajectories in order to predict
%  selectivity since they undershoot the MD MSD by a consistent number and 
%  the ratios will stay the same.

  % BJC2: the numbers and choices of states in this paragraph will change, but I think we
  % can still use this framework to study mechanisms.

%  \subsubsection*{Relating Clusters to Mechanisms}
%  
%  The parameters of each cluster can be related to dominant solute-membrane interactions.
%  \begin{itemize}
%    \item Based on our previous work, we know that hydrogen bonding, sodium ion 
%    association and radial location within the membrane pores will have a significant
%    influence on transport properties. 
%    \item Even without our previous work, it would still be easy to speculate that these 
%    interactions might occur and could therefore be used to help understand the 
%    states identified by the IHMM.
%    \item We have also incorporated membrane density as a physical metric because we 
%    know the membrane has an inhomogeneous structure.
%  \end{itemize}

%  \textit{Learning from individual trajectories}: In Figure~\ref{fig:mechanism_map}, 
%  we demonstrate the wide range of behavior exhibited by a representative 2D methanol
%  trajectory.
%%BJC2: reducing duplication
%%  The coordinates
%%  are colored according to the cluster with which methanol's dynamics are most consistent,
%%  which implies that this particular methanol trajectory exhibits eight distinct dynamical 
%%  behaviors (in order of their first appearance, they are colored dark yellow, red, light
%%  blue, light green, orange, blue, teal and lime green).
%  The solute time series are coupled with color coded bars which describe the solute's
%  local environment and physical interactions that it undergoes at each trajectory frame. 
%%  From the top, they describe 
%%  whether the solute is associated with sodium, whether the solute is hydrogen bonding
%%  with the monomer as well as to what part of the monomer it is hydrogen bonding, and 
%%  finally, the local number density surrounding the solute. 
%
%  %BJC2: the exact figure will change, but most of what I say shouldn't
%  \begin{figure}
%  \centering
%  \begin{subfigure}{0.75\textwidth}
%  \includegraphics[width=\textwidth]{mechanism_map.png}
%  \caption{}
%  \end{subfigure}
%  \begin{subfigure}{0.24\textwidth}
%  \includegraphics[width=\textwidth]{monomer_oxygens.png}
%  \caption{}
%  \end{subfigure}
%  %MRS: a lot of this caption test is duplicated from the main text. I would suggest reducing the duplication.
%  %BJC: you think better to have more detail in text or in caption?
%  %MRS2: at a certain point, it should be in the text.  In the caption, I would just state what the point of the plot is, 
%  % and define the colors/data/etc.  But full explanation would be better in the text. Captions should be scannable quickly,
%  % based on how people process papers, and this should be significantly shorter for readability. 
% \caption{(a) We can learn a significant amount of detail about solute motion by viewing 
%  solute trajectories color-coded according to their clustered dynamical behavior alongside
%  plots of the dominant molecular interactions and descriptions of the membrane's local 
%  environment. In the plot above, we show the radial and axial coordinates of a single
%  methanol molecule over the course of the equilibrated portion of its MD simulation.
%  Disjoint segments of the same color imply the solute behaves similarly at different 
%  points of the trajectory because their VAR parameters place them in the same cluster. 
%  There are eight distinct dynamical modes exhibited by this trajectory. In order
%  of their first appearance, they are colored dark yellow, red, light blue, light green,
